An investigation into using scagnostics to find interesting projections of high-dimensional data

### 2018-10-30 scagnostics.Rmd
Some text updates in the summary.

Our results can be summarised by evaluating and comparing the advantages and disadvanteges of each index function according to the criteria presented above. An overview is given in Table \ref{tab:summary}. The first three entries are from the scagnostics family. The advantages are good scaling of computing time with sample size and that the set of index functions covers general signatures (the same is however not true for each single index viewed on its own). On the other hand we have found that they are typically too noisy making optimisation of the function challenging. In particular the stringy index cannot be used with current optimisers. The index functions from @mbgraphic share similar behaviour as they are both found to evolve fairly smoothly between interpolated projections, but have been shown to be strongly rotation dependent. The splines2D index is fast to calculate and found to be the most efficient method for identifying functional dependence. We have only considered the non-binned implementation of dcor2D which is slow to calculate on large datasets. While this index is in principle general, we have found it to be too sensitive to noise to be applicable in most realistic examples. Finally the MINE indexes are found to be smooth only for large datasets, with some rotation dependence observed for the TIC index. While these indexes are slow to compute especially on large datasets, we have found them to be the most general, allowing good discrimination even with noisy structures.

\begin{table}
\begin{center}
\caption{Summary}
\label{tab:summary}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
Index & smooth & squintability & flexible & rotation invariant & speed \\
\hline
Convex & ! & ! & ! & ! & \checkmark \\
Skinny & ! & !& ! & ! & \checkmark \\
Stringy & $\times$ & $\times$ & ! & ! & \checkmark \\
\hline
splines2D & \checkmark & \checkmark & ! & $\times$ & \checkmark \\
dcor2D & \checkmark & \checkmark & ! & $\times$ & ! \\
\hline
MINE & ! & \checkmark  & \checkmark & ! & ! \\
\hline
\end{tabular}
\end{center}
\end{table}

>>>>

Our results can be summarised by evaluating and comparing the advantages and disadvanteges of each index function according to the criteria presented above. An overview is given in Table \ref{tab:summary}. The first three entries are from the scagnostics family. The advantages are good scaling of computing time with sample size and that the set of index functions covers general signatures, making the set flexible in applications (the same is however not true for each single index viewed on its own). On the other hand we have found that they are typically too noisy making optimisation of the function challenging, which also results in poor squinatbility. In particular the stringy index cannot be used with current optimisers. The index functions from @mbgraphic share similar behaviour as they are both found to evolve fairly smoothly between interpolated projections, but have been shown to be strongly rotation dependent. The splines2D index is fast to calculate and found to be the most efficient method for identifying functional dependence. We have only considered the non-binned implementation of dcor2D which is slow to calculate on large datasets. While this index is in principle general, we have found it to be too sensitive to noise to be applicable in most realistic examples, therefore we consider flexibility limited. Finally the MINE indexes are found to be smooth only for large datasets, with some rotation dependence observed for the TIC index. While these indexes are slow to compute especially on large datasets, we have found them to be the most general, allowing good discrimination even with noisy structures. In addition we observed good squinatbility for all three examples considered, given large datasets. Results in Section \ref{sec:params} suggest that tuning of the MINE parameters may result in improved behaviour for smaller datasets.

\begin{table}
\begin{center}
\caption{Summary of findings, showing to what extend the considered index functions pass the criteria for a good PPI. Checkmark symbols good behaviour, exclamation mark symbols some issues while cross symbols failure on the corresponding criteria. Each index has particular strengths and drawbacks and selection must be guided by the considered example, see text for details.}
\label{tab:summary}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
Index & smooth & squintability & flexible & rotation invariant & speed \\
\hline
Convex & ! & ! & ! & ! & \checkmark \\
Skinny & ! & !& ! & ! & \checkmark \\
Stringy & $\times$ & $\times$ & ! & ! & \checkmark \\
\hline
splines2D & \checkmark & \checkmark & ! & $\times$ & \checkmark \\
dcor2D & \checkmark & \checkmark & ! & $\times$ & ! \\
\hline
MINE & ! & \checkmark  & \checkmark & ! & ! \\
\hline
\end{tabular}
\end{center}
\end{table}
